### 2017-05-08 paper2.tex

\subsection{Seasonality in sUHI magnitudes}

@@

Figure \ref{heatcluhi} shows seasonal variations in diurnal patterns of mean monthly clUHI calculated for the same study period as Figure \ref{heatsuhi}. clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset. Daytime clUHI does not vary significantly across seasons.

@@

@@

@@

@@

Figures \ref{meteo_kdown}, \ref{meteo_hum}, \ref{meteo_wv}, and \ref{meteo_atuhi} show hemispherical sUHI magnitudes as a function of incoming solar radiation, water vapor content, wind velocity, and clUHI measured over the BUBBLE campaign. Winter hours are omitted in Figures \ref{meteo_hum}, and \ref{meteo_wv}, and \ref{meteo_atuhi} as sUHI is small and displays minimal variation in winter. Both winter and nighttime hours are omitted in Figure \ref{meteo_kdown} to examine daytime sUHI under clear sky and clouded sky conditions.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 18 \si{\mega\joule\per\meter\squared}) and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 18 \si{\mega\joule\per\meter\squared})}
	\label{meteo_kdown}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
	\caption{sUHI magnitude versus rural wind velocity measured at approximately 10 \si{\meter} above ground at the Lange Erlen site.}
	\label{meteo_wv}
\end{figure}

@@

\section{Discussion}
\citet{Oke2017} presents a generalized conceptual model for sUHI development for multiple representations of the surface. However, the results in \citet{Oke2017} are based on simulations of an idealized urban area without trees, complex non-orthogonal surface geometry, or sloped facets and have not been supported observationally. The results in this paper provides an ideal platform to test conceptual understanding of sUHI development and to understand the diurnal and seasonal nature of sUHI.

@@

sUHI development is the result of differences in urban and rural heating and cooling rates. A variety of factors modify urban heating and cooling rates relative to non-urban spaces, most notably urban-rural contrasts in sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts). These differences provide the genesis for diurnal sUHI development. Rural surfaces have more open (less obstructed) sky view factors, and respond quickly to insolation a low solar angles in the early morning hours. Shading by canyon features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunset. This is observed in Basel across all representations of urban T\textsubscript{surf} at similar magnitudes. Through the late morning hours, urban heating rates increase relative to rural rates as the solar angle increases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up. Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces. After the initial spike in rural heating in the morning hours, urban heating rates overtake rural and the sUHI slowly grows through midday. In the hours after midday, the solar angle decreases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to cool. sUHI reaches a peak at approximately two to three hours hours after solar noon.

In the afternoon hours, both urban and rural surfaces transition to cooling, with approximately equal rates. The late afternoon and evening decline in sUHI is strongly controlled by urban surface structure. Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets. These two factors control whether sUHI grows or shrinks in the hours after sUHI\textsubscript{max}. In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI. The opposite is true of urban areas with obstructed sky view factors, and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening. For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping.

After sunset, rural cooling rates spike relative to urban, as an unobstructed rural sky-view-factor allows for strong radiative cooling of the surface, leading to a slight jump in sUHI.  However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates. All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics. This is evident in characterization of sUHI in Basel, where the two rural sites entail different rural cooling rates, particularly in the hours just after sunset, and different patterns of post-sunset sUHI development. Although similar in terms of surface coverage and site geometry, BUBBLE rural sites are subject to different after sunset air masses that strongly effect sUHI (and clUHI) development. Hilly terrain surrounding the Lange Erlen site results in cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days. This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses. The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site. This makes clear the need for proper characterization of the radiative and thermal properties, surface geometry and coverage of rural sites in sUHI analysis, as well as the surrounding topography to ensure characterizations of sUHI development are free of confounding influence. After initial nighttime divergence, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.

@@

Hemispherical sUHI in Basel displays significant seasonality, both in terms of its diurnal pattern and overall magnitude. Seasonal variations in sUHI are most evident during the daytime hours, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development. Nighttime sUHI is consistent across seasons. Peak seasonal sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense, forcing significant contrasts in urban-rural heating/cooling rates. In winter, daytime sUHI is suppressed from a lower solar input and frequent cloud cover. These reduce urban-rural contrasts in heating/cooling rates and weaken wintertime sUHI. In Basel, wintertime sUHI has a small peak shortly after sunset, when cold air drainage amplifies rural cooling rates and increases sUHI. This, the small after sunset peak in sUHI present in both summer and winter observations is likely not generalizable and is specific to the Lange Erlen site.

These results are supported by similar seasonal patterns of hemispherical sUHI observed in Vancouver, where seasonal sUHI\textsubscript{max} (sUHI\textsubscript{min}) is observed near the summer (winter) solstice. Observations of sUHI in Vancouver show similar seasonal patterns of development to those in Basel, with sUHI with a larger diurnal sUHI amplitude in summer months and suppressed sUHI development in winter. Cooling rates at the Westham island rural site are not affected by cold air drainage and, as a result, sUHI observed in Vancouver does not display two diurnal peaks in summer nor a coherent peak after sunset in winter months.

Although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns. In contrast to sUHI, clUHI in Basel is greatest just after sunset following closely to sunset/sunrise timing. Peak clUHI occurs just after sunset across seasons, differing in terms of magnitude but not time of day, with a larger peak clUHI in summer\footnote{As is the case with sUHI observed in Basel, clUHI magnitudes are effected by cold air drainage at the Lange Erlen site and are slightly higher than expected \citep{Runnalls2000}.}. Daytime clUHI is small and displays little seasonal variation. Thus, a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.

@@

Sensor-surface geometry exerts significant influence on observed sUHI magnitudes. By undersampling vertical and sloped urban surfaces and oversampling low thermal admittance fabric materials, sUHI\textsubscript{nadir} overestimates mean afternoon sUHI\textsubscript{comp} by 3.41 \si{\kelvin} and up to 5.96 \si{\kelvin} under clear sky “satellite friendly” conditions. Thus, peak sUHI\textsubscript{nadir} overestimates the 'true' sUHI\textsubscript{comp} by nearly double under clear sky conditions. As clear sky conditions are the only conditions under which satellite TIR remote sensors can operate, and make up a large fraction of the urban T\textsubscript{surf} literature, these biases are pervasive and cannot be ignored. In contrast, by sampling the urban surface in 3-dimensions, sUHI\textsubscript{hem} yields a much smaller mean daytime overestimation (0.54 \si{\kelvin}), particularly under clear sky conditions (0.93 \si{\kelvin}), providing a significantly more representative depiction of urban T\textsubscript{comp} and sUHI\textsubscript{comp}, in addition to allowing for time-continuous analysis. Cloud cover suppresses the influence of sensor-surface geometry on remote sensed sUHI by reducing microscale geometric and spatiotemporal contrasts in urban T\textsubscript{surf}, resulting in less facet-scale heterogeneity in urban T\textsubscript{surf} and reducing the overall magnitude of sUHI for all definitions.

@@

sUHI is highly dependent on meteorological conditions, which account for much of the day to day variation sUHI magnitudes. Conditions that foster the largest microscale contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes. Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show similarly reduced sUHI magnitudes. As such, incoming shortwave radiation and wind velocity display positive and negative relationships with sUHI magnitudes respectively. Under clear sky conditions, strong solar heating of the surface combined with differences in urban and rural thermal properties and sunlit shading geometries fosters the development of large daytime urban-rural contrasts in T\textsubscript{surf} and T\textsubscript{air} and the development of strong sUHI by day and clUHI by night. Thus, sUHI\textsubscript{max} coincides with insolation\textsubscript{max}. However, significanyUnder cloudy conditions, decreased direct solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in a reduced sUHI. $\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI, owing to its dependence on solar insolation.

As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself. Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry - shown in Figure \ref{humtest}. As water vapor content increases from zero, variation in spectral atmospheric transmittance decreases. Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing longwave radiation available for heating the canopy layer air volume through absorption and for heating of the surface by atmospheric emission. As a result, water vapor content has little effect on sUHI. Wind velocity displays a strong negative relationship with sUHI. sUHI is greatest under calm winds when near-surface mixing is minimal, fostering strong microscale and urban-rural contrasts in T\textsubscript{surf}. As wind velocity increases, sUHI is suppressed from increased turbulent heat losses. During the nighttime hours, increased wind velocities destroy the surface based thermal inversion often observed over rural terrain. This further limits nighttime cooling rates relative to urban and decreases sUHI magnitudes. Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.

\begin{table}[H]
	\centering
	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
	\label{meteo_cont}
	\begin{tabular}{lc}
		\toprule 
		An increase in $x$, has a/an... & $y$ effect on sUHI magnitude. \\
		\midrule
		$\Delta$T\textsubscript{surf - air} &\textit{increasing} \\
		&\\
		T\textsubscript{surf} & \textit{increasing} \\
		&\\
		T\textsubscript{air} & \textit{increasing} \\
		&\\
		Incoming solar radiation & \textit{increasing}  \\
		& \\
		Water vapor content & \textit{negligible} \\
		&\\
		Wind velocity &\textit{decreasing} \\
		&\\
		clUHI &\textit{increasing} \\
		\bottomrule
	\end{tabular} 
\end{table}

clUHI and sUHI develop at different times of day and, thus, have a complex diurnal relationship. By day, variation in clUHI is minimal and shows a weak positive correlation with sUHI. By night, clUHI shows greater variation and a stronger, more positive relationship with sUHI. The relationship between clUHI and sUHI is most evident on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest. Under these conditions sUHI and clUHI  show similar development in the hours after sunset.

@@

This chapter presents the results of a time-continuous eight month climatological analysis of the surface urban heat island in Basel, Switzerland derived using the method presented in Chapter \ref{paper1}. Hemispherical sUHI magnitudes display significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong microscale and urban-rural contrasts in T\textsubscript{surf}. Seasonality in late afternoon sUHI is largely dependent on incoming solar radiation, which displays a strong positive relationship with sUHI magnitudes. In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development resulting in less day-to-day variation in sUHI, and lower overall sUHI magnitudes.

Remote sensed sUHI is highly dependent on sensor-surface-sun geometry. This results in significant variation in urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes. Thus, traditional methods for thermal remote sensing of urban areas do not capture the true temporal and geometric nature of the urban effect on land T\textsubscript{surf}. Urban effective anisotropy results in large geometric biases in remote sensing of urban areas as traditional satellite and aerial remote sensors sample only a fraction of the urban surface. These geometric biases manifest not only in deviation between observed sUHI magnitudes and the 'true' complete sUHI, but also impact characterizations of diurnal and seasonal patterns of sUHI and urban T\textsubscript{surf} by changing the time of day of sUHI\textsubscript{max}. Geometric biases in thermal remote sensing of urban area are compounded by myriad temporal biases across a range of temporal scales, as traditional methods for urban thermal remote sensing are not time-continuous and only operable under clear sky conditions. Both geometric and temporal biases affect the vast majority of sUHI study, most of which is performed using only a small number of satellite remote sensors, to which these biases are inherent and largely unavoidable.

>>>>

\subsection{Diurnal and seasonal variability in sUHI magnitudes}

@@

Figure \ref{heatcluhi} shows seasonal variation in diurnal patterns of mean monthly clUHI calculated for the same study period as Figure \ref{heatsuhi}. clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset. Daytime clUHI does not vary significantly across seasons.

@@

Figure \ref{heat_uhi_diff} shows seasonal variation in the difference between mean monthly hemispherical sUHI\textsubscript{norm} and clUHi\textsubscript{norm} for the BUBBLE campaign. By restricting UHI values to a range between one and zero, diurnal patterns of sUHI and clUHI can be compared directly. Comparison of the two in summer shows a peak sUHI approximately two hours after solar noon (when clUHI is small) and peak clUHI approximately two hours after sunset (when sUHI is small). Values near zero in the winder months indicate little variation in both sUHI and clUHI.

@@

Figure \ref{vanc_suhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the EPiCC campaign. The sUHI in Vancouver shows similar seasonal and diurnal patterns to the sUHI in Basel, but does not persist as far into the late evening hours.

@@

Figure \ref{bxsUHI_season_bubble} shows hemispherical sUHI magnitudes for the eight month BUBBLE campaign and binned for winter (December through February) and summer (April through July) seasons. Diurnal variability in sUHI magnitudes is concentrated in summer months, as well as seasonal sUHI\textsubscript{max}.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_season_bubble}
	\caption{Hemispherical sUHI magnitudes for the eight month BUBBLE climatology (top) and binned for winter (middle) and summer (bottom) seasons. Grey shading indicates nighttime hours averaged over the bin interval.}
	\label{bxsUHI_season_bubble}
\end{figure}

@@

Figure \ref{meteo_wv_sol} panel (a) shows mean and maximum sUHI magnitudes as a function of solar radiant exposure over each day of the summer months of the BUBBLE campaign. Point shading indicates the month (white is April and black is July). A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}. Panel (b) of Figure \ref{meteo_wv_sol} shows mean and maximum hemispherical sUHI magnitudes as a function of mean wind velocity for the duration of the eight month BUBBLE climatology. Colors indicate solar radiant exposure accumulated over each day. Similar results are observed when rural winds are substituted.

Figures \ref{meteo_hum} and \ref{meteo_atuhi} show hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured over the BUBBLE campaign. Winter hours are omitted in both as sUHI is small and displays minimal variation in winter. Jamie: still working on getting color palettes for these. Will merge into one figure.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv_sol}
	\caption{(a) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July versus integrated solar radiant exposure measured over the same 24 \si{\hour} period. Shading indicates month of the year, with dark points indicating later months. (b) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day in the eight month climatology versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site. Coloring indicates integrated solar radiant exposure over each day.}
	\label{meteo_wv_sol}
\end{figure}

@@

@@

\section{Discussion}
\citet{Oke2017} presents a generalized conceptual model for sUHI development for multiple representations of the surface. However, the results in \citet{Oke2017} are based on simulations of an idealized urban area without trees, complex non-orthogonal surface geometry, or sloped facets and have not been supported observationally. The results in this chapter provide an ideal platform to test conceptual understanding of sUHI development, to understand the diurnal and seasonal nature of sUHI and the meteorological factors that drive its development.

@@

sUHI development is the result of urban modification to surface geometry and thermal and radiative properties, most notably sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts). These differences alter urban heating/cooling rates relative to non-urban spaces and provide the genesis for diurnal sUHI development. In the early morning hours, rural surfaces respond quickly to insolation at low solar angles, and rural T\textsubscript{surf} rapidly increases. In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise. This is observed in Basel across all representations of urban T\textsubscript{surf} at similar magnitudes. Through the late morning hours, urban heating rates increase relative to rural rates as the solar angle increases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up. Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces. After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two to three hours hours after solar noon. In the hours after UHI\textsubscript{max}, the solar angle decreases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.

In the afternoon hours, both urban and rural surfaces transition to cooling, this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure. Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets. These two factors control whether sUHI grows or shrinks in the hours after sUHI\textsubscript{max}. In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI. The opposite is true of urban areas with restricted sky view factors and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening. For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping. The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel. A nadir view, by virtue of oversampling rapidly cooling rooftops, observes a strong evening decline in sUHI. This is roughly equivalent to sUHI calculated from an urban site with low canyon height to width ratio and a relatively unobstructed sky view factor. In contrast, by sampling vertical facets that cool more gradually, this decline is less evident in sUHI calculated from a complete view of Basel T\textsubscript{surf}. Thus, as expected, remote sensed sUHI analysis is also directionally dependent and subject to anisotropic effects when viewed from a narrow-FOV sensor. These effects not only affect the magnitude of remote sensed sUHI, but may also alter perceived diurnal patterns of sUHI.

After sunset, the unobstructed rural sky-view-factor allows for strong radiative cooling of the surface and rural cooling rates spike relative to urban, leading to a slight jump in sUHI. However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates. All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics. This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel. Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days. This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses. The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site. Following site dependent effects on sUHI in the hours after sunset, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.

% This makes clear the need for proper characterization of the radiative and thermal properties, surface geometry and coverage of rural sites in sUHI analysis, as well as the surrounding topography to ensure characterizations of sUHI development are free of confounding influence.

@@

Hemispherical sUHI in both Basel and Vancouver displays significant seasonality, both in terms of its diurnal pattern and overall magnitude. Seasonal variations in sUHI are most evident during the daytime hours, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development. Mean monthly sUHI\textsubscript{max} is observed near the summer solstice in both study areas, when days are longest and insolation on clear sky days is most intense, forcing significant contrasts in urban-rural heating/cooling rates and the development of a strong sUHI. In winter, lower overall solar input and frequent cloud cover suppress the potential for urban-rural contrasts in heating/cooling rates and weakens wintertime sUHI development. In Basel, wintertime sUHI has a small peak shortly after sunset, when cold air drainage anomalously amplifies rural cooling rates, increasing the sUHI. This small, after-sunset peak in sUHI observed in both summer and winter observations is specific to the Lange Erlen site and is not generalizable. Cooling rates at the Vancouver Westham island rural site are not affected by cold air drainage and, as a result, sUHI does not display two diurnal peaks in summer nor a coherent peak after sunset in winter months. Thus, observations of seasonal variations in hemispherical sUHI in Vancouver are likely highly generalizable to other mid-latitude cities with distinct winter and summer seasons.

The sUHI and the clUHI have similar micro meteorological genesis, climatic controls, and display similar seasonal variability, albeit concentrated at different times of the day. This is evident when comparing normalized mean monthly hemispherical sUHI and clUHI - where the range of both types of sUHI is restricted to values between zero and one. In winter, low solar input and frequent cloud cover inhibit variability in both sUHI and clUHI and both are of similar normalized magnitudes. In summer, larger solar input fosters contrasts in urban-rural heating and cooling rates and both clUHI and sUHI are most variable, with peak sUHI in the early afternoon clUHI is smallest, and peak clUHI in the hours after sunset when sUHI is relatively low. These results suggest the following generalization: when conditions that foster the formation of strong urban-rural contrasts in T\textsubscript{air} and T\textsubscript{surf} are present (most notably a large solar input), sUHI peaks approximately two hours after solar noon and clUHI peaks approximately two hours after sunset. Thus, a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.

@@

Sensor-surface geometry exerts significant influence on observed sUHI magnitudes. 2-dimensional, complete, and hemispherical observations of sUHI not only yield different magnitudes but also different diurnal sUHI regimes and peak timings. Variations in observed sUHI across different sensor sampling regimes are largest under clear sky, hot conditions, coinciding with typical seasonal and diurnal sUHI\textsubscript{max}. In general, as urban T\textsubscript{surf} is most often viewed from the top down, remote sensors preferentially sample horizontally oriented facets with unobstructed sky view factors and undersample or ignore vertical and sloped facets with more obstructed sky view factors. Over/undersampling of the urban surface changes both the magnitude and character of remote sensed urban T\textsubscript{surf} and sUHI.

In an analysis of Basel sUHI from complete, nadir, and hemispherical representations of the surface, significant variation in remote sensed sUHI was observed across sensor sampling regimes at all times of the day. Variations in observed sUHI are most evident on clear sky, hot days in the hours after solar noon, when microscale spatiotemporal contrasts in urban T\textsubscript{surf} are largest. Cloud cover suppresses the influence of sensor-surface geometry on remote sensed sUHI by reducing microscale geometric and spatiotemporal contrasts in urban T\textsubscript{surf}, resulting in less facet-scale heterogeneity in urban T\textsubscript{surf} and reducing the overall magnitude of sUHI for all definitions.

By undersampling vertical and sloped urban surfaces and oversampling low thermal admittance materials, sUHI\textsubscript{nadir} overestimates mean afternoon sUHI\textsubscript{comp} by approximately 3.5 \si{\kelvin} and up to 6.0 \si{\kelvin} under clear sky “satellite friendly” conditions. This constitutes a significant bias in remote sensed sUHI\textsubscript{max} as sUHI\textsubscript{nadir} overestimates the 'true' sUHI\textsubscript{comp} by nearly double under clear sky conditions. As clear sky conditions are the only conditions under which satellite TIR remote sensors can operate, and make up a large fraction of the urban T\textsubscript{surf} literature, these biases are pervasive and cannot be ignored. In contrast to sUHI\textsubscript{nadir}, sUHI\textsubscript{hem} samples the urban surface in 3-dimensions ("seeing" a wider range of facet types and orientations) and yields a much smaller mean daytime overestimation (0.5 \si{\kelvin}), particularly under clear sky conditions (1.0 \si{\kelvin}), providing a significantly more representative depiction of urban T\textsubscript{comp} and sUHI\textsubscript{comp}, in addition to allowing for time-continuous analysis.

@@

sUHI is highly dependent on meteorological conditions, which account for much of the day to day variation sUHI magnitudes. Conditions that foster the largest microscale spatiotemporal contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes. Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show reduced sUHI magnitudes. As such, sUHI development is strongly tied to integrated solar radiant exposure accumulated over a day and wind velocity, which display positive and negative relationships with sUHI magnitudes respectively. Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.

The relationship between solar radiation and sUHI is relatively simple but easily misinterpreted. Surfaces (both urban and otherwise) change in temperature based on accumulated radiative or convective gains or losses - as such, instantaneous changes in insolation have little effect on sUHI magnitudes. Thus, to say that sUHI is a function of measured incoming solar radiation (in \si{\watt\per\square\meter}) does not accurately characterize the relationship and produces a significant amount of unrepresentative scatter. Strong accumulated solar heating of the surface combined with differences in urban and rural thermal properties and sunlit shading geometries fosters the development of large daytime urban-rural contrasts in T\textsubscript{surf} and T\textsubscript{air} and the development of strong sUHI by day and clUHI by night. Thus, sUHI\textsubscript{max} often coincides with insolation\textsubscript{max}. Under cloudy conditions, reduced accumulated solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in reduced sUHI magnitudes. Owing to its dependence on accumulated solar heating, $\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI magnitudes.

Wind velocity has a similarly nuanced relationship with sUHI. Winds affect surface heating and cooling rates by facilitating convective cooling (or, under less common circumstances, convective heating). In general, as wind velocity increases, convective heat losses increase and T\textsubscript{surf} decreases. This suppresses microscale spatiotemporal variations in T\textsubscript{surf} as hot sunlit surfaces experience greater rates of convective cooling under high wind conditions than cooler shaded surfaces. This reduces overall urban and rural T\textsubscript{surf} and inhibits sUHI development. Nighttime sUHI is also affected by wind velocity as the formation of a near surface thermal inversion, which increases radiative cooling in non-urban areas, requires calm winds. As such, increased nighttime wind velocity decreases nocturnal sUHI.

As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself. Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry - shown in Figure \ref{humtest}. As water vapor content increases from zero, variation in spectral atmospheric transmittance decreases. Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing longwave radiation available for heating the canopy layer air volume through absorption and for heating of the surface by atmospheric emission. As a result, water vapor content has little effect on sUHI.

\begin{table}[H]
	\centering
	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
	\label{meteo_cont}
	\begin{tabular}{lc}
		\toprule 
		An increase in $x$, has a/an... & $y$ effect on sUHI magnitude. \\
		\midrule
		$\Delta$T\textsubscript{surf - air}, \si{\kelvin} &\textit{increasing} \\
		&\\
		T\textsubscript{surf}, \si{\degreeCelsius}  & \textit{increasing} \\
		&\\
		T\textsubscript{air}, \si{\degreeCelsius} & \textit{increasing} \\
		&\\
		Insolation, \si{\mega\joule\per\meter\squared\per\day} & \textit{increasing}  \\
		& \\
		Water vapor content, \si{\gram\per\meter\squared} & \textit{negligible} \\
		&\\
		Wind velocity, \si{\meter\per\second} &\textit{decreasing} \\
		&\\
		clUHI, \si{\kelvin} &\textit{increasing} \\
		\bottomrule
	\end{tabular} 
\end{table}

clUHI and sUHI develop at different times of day and, thus, have a complex diurnal relationship. By day, variation in clUHI is minimal and shows weak positive correlation with sUHI. Occasionally, under clear sky conditions following hot days, strong evening radiative cooling forces clUHI development slightly before sunset. This causes a slight bifurcation in clUHI versus sUHI. By night, clUHI shows greater variation and a stronger, more positive relationship with sUHI, most evident on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest.

@@

This chapter presents the results of an analysis of two time-continuous long-term climatologies of the surface urban heat island effect in Basel, Switzerland and Vancouver, Canada derived using the method presented in Chapter \ref{paper1}. Hemispherical sUHI magnitudes display significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong urban-rural contrasts in T\textsubscript{surf}. Seasonality in late afternoon sUHI is largely dependent on accumulated radiative exposure over the day which displays a strong positive relationship with sUHI magnitudes. In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development resulting in less variability in sUHI, and lower overall sUHI magnitudes.

Remote sensed sUHI is highly dependent on sensor-surface-sun geometry. This results in significant variation in the character of urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes. Thus, traditional methods for thermal remote sensing of urban areas do not capture the true temporal and geometric nature of the urban effect on land T\textsubscript{surf}. Urban effective anisotropy results in large geometric biases in remote sensing of urban areas as traditional satellite and aerial remote sensors sample only a fraction of the urban surface. These geometric biases manifest not only in deviation between observed sUHI magnitudes and the 'true' complete sUHI, but also impact characterizations of diurnal and seasonal patterns of sUHI and urban T\textsubscript{surf}. Geometric biases in thermal remote sensing of urban area are compounded by myriad temporal biases across a range of temporal scales, as traditional methods for urban thermal remote sensing are not time-continuous and only operable under clear sky conditions. Both geometric and temporal biases affect the vast majority of sUHI study, most of which is performed using only a small number of satellite remote sensors, to which these biases are inherent.
